
#### Introduction
Data given is a de-identified Excel spreadsheet. Assignment is to perform the appropriate analysis to forecast several series for 140 periods.We are having 1622 periods for our analysis. 

Loading al the required libraries.

```{r,message=FALSE,warnings=FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(gridExtra)
```

Read in data, which is just the Excel sheet saved as a CSV.

```{r}
data <- read.csv("Project1data.csv",header=TRUE)
data$SeriesInd <- as.Date(data$SeriesInd,origin = "1899-12-30")
project_in_df <- data
```
As part of data cleaning lets remove forecast cells for now


```{r}
project_in_df <- data[1:9732,]
```

Lets have a look at data
```{r}
summary(project_in_df)
```

From the above results can see missing values in all the variables,we need to impute NA's with some value.
Var01 has 14,Var02 has 2,Var03 has 26,Var05 has 26 and Var07 has 26 missing values.

Here we will be imputing these values with  the next value in sequence, e.g., if row 1600 is NA for Var01, take value from row 1601
hopefully, since these are stock values, the value from the next day for the few missing values we have, should be close enough

```{r}
project_in_df[rowSums(is.na(project_in_df))>0,]
while(nrow(project_in_df[rowSums(is.na(project_in_df))>0,]) > 0)
{           
    
  project_in_df <- transmute(project_in_df, 
                      SeriesInd = SeriesInd,group,
                      Var01 = if_else(is.na(Var01), lead(Var01), Var01),
                      Var02 = if_else(is.na(Var02), lead(Var02), Var02),
                      Var03 = if_else(is.na(Var03), lead(Var03), Var03),
                      Var05 = if_else(is.na(Var05), lead(Var05), Var05),
                      Var07 = if_else(is.na(Var07), lead(Var07), Var07))
    print(project_in_df[rowSums(is.na(project_in_df))>0,])
}
summary(project_in_df)
```

From the above summary results we can see all Null values are replaced


# Create separate dataframes for each group
```{r}
group_S01_df = filter(project_in_df, group == 'S01')
group_S02_df = filter(project_in_df, group == 'S02')
group_S03_df = filter(project_in_df, group == 'S03')
group_S04_df = filter(project_in_df, group == 'S04')
group_S05_df = filter(project_in_df, group == 'S05')
group_S06_df = filter(project_in_df, group == 'S06')
```


Checking for the missing values inour dataset

```{r}
group_S01_df[rowSums(is.na(group_S01_df))>0,]
group_S02_df[rowSums(is.na(group_S02_df))>0,]
group_S03_df[rowSums(is.na(group_S03_df))>0,]
group_S04_df[rowSums(is.na(group_S04_df))>0,]
group_S05_df[rowSums(is.na(group_S05_df))>0,]
group_S06_df[rowSums(is.na(group_S06_df))>0,]
```




```{r}
# Select relevant columns for each group
group_S01_df = select (group_S01_df, matches("SeriesInd|datetime|group|Var01|Var02"))
group_S02_df = select (group_S02_df, matches("SeriesInd|datetime|group|Var02|Var03"))
group_S03_df = select (group_S03_df, matches("SeriesInd|datetime|group|Var05|Var07"))
group_S04_df = select (group_S04_df, matches("SeriesInd|datetime|group|Var01|Var02"))
group_S05_df = select (group_S05_df, matches("SeriesInd|datetime|group|Var02|Var03"))
group_S06_df = select (group_S06_df, matches("SeriesInd|datetime|group|Var05|Var07"))
# Check number of rows
print (c(nrow(group_S01_df), nrow(group_S02_df), nrow(group_S03_df), nrow(group_S04_df), nrow(group_S05_df), nrow(group_S06_df)))
# Verify dataframes
head(project_in_df, 20)
head(group_S01_df, 20)
head(group_S02_df, 20)
head(group_S03_df, 20)
head(group_S04_df, 20)
head(group_S05_df, 20)
head(group_S06_df, 20)
```


#### <span style="color:red"><b><u>Exploratory Data Analysis</u></b></span>


```{r}
p1 <- ggplot(project_in_df, aes(SeriesInd, y = Var02, color = variable)) + 
    geom_line(aes(y = Var02, col = "Var02"))

p2 <- ggplot(group_S01_df, aes(SeriesInd, y = Var01, color = variable)) + 
    ggtitle("S01") + 
    geom_line(aes(y = Var01, col = "Var01"))

p3 <- ggplot(group_S02_df, aes(SeriesInd, y = Var03,color = variable))+
    ggtitle("S02") + 
    geom_line(aes(y = Var03, col = "Var03"))

p4 <- ggplot(group_S03_df, aes(SeriesInd, y = Var05, color = variable)) + 
    geom_line(aes(y = Var05, col = "Var05")) +
    ggtitle("S03") 

p4 <- ggplot(group_S03_df, aes(SeriesInd, y = Var07, color = variable)) + 
    geom_line(aes(y = Var07, col = "Var07")) +
    ggtitle("S03") 
  
p5 <- ggplot(group_S04_df, aes(SeriesInd, y = Var01, color = variable)) + 
    geom_line(aes(y = Var01, col = "Var01")) + 
    ggtitle("S04") 

p6 <- ggplot(group_S05_df, aes(SeriesInd, y = Var03, color = variable)) + 
   ggtitle("S05") +
    geom_line(aes(y = Var03, col = "Var03"))

p7 <- ggplot(group_S06_df, aes(SeriesInd, y = Var05, color = variable)) + 
    geom_line(aes(y = Var05, col = "Var05")) +
    ggtitle("S06")

p8 <- ggplot(group_S06_df, aes(SeriesInd, y = Var07, color = variable)) + 
     ggtitle("S06") +
    geom_line(aes(y = Var07, col = "Var07"))
grid.arrange(p1, p2, nrow = 1)
grid.arrange(p3, p4, nrow = 1)
grid.arrange(p5, p6, nrow = 1)
grid.arrange(p7,p8, nrow = 1)
```

Comparing to other variable Var02 seems to be noisy than any other variable and having outliers.
These outliers needs to be fixed before producing forecasts
S03 and S06 variables Var05 and Var07 seems to be quite similar
Also we see in S02 - Var03 and S06-Var05 and Var07 plot some outlier values that also needs to be fixed before forecasting.
We can also observe some seasonality and trend pattern in Var01,Var03,Var05 and var07


Removing outliers
We are using here IQR to fix the outliers in our data
For missing values that lie outside the 1.5*IQR limits, we are capping it by replacing those observations outside the lower limit with the value of 5th %ile and those that lie above the upper limit, with the value of 95th %ile.
```{r}
Remove_Outlier <- function(x){
   qnt <- quantile(x, probs=c(.25, .75), na.rm = T)
   caps <- quantile(x, probs=c(.05, .95), na.rm = T)
   H <- 1.5 * IQR(x, na.rm = T)
   x[x < (qnt[1] - H)] <- caps[1]
   x[x > (qnt[2] + H)] <- caps[2]
   return(x)
}

group_S01_df$Var01=Remove_Outlier(group_S01_df$Var01)
group_S02_df$Var03=Remove_Outlier(group_S02_df$Var03)
project_in_df$Var02=Remove_Outlier(project_in_df$Var02)
group_S04_df$Var01=Remove_Outlier(group_S04_df$Var01)
group_S05_df$Var03=Remove_Outlier(group_S05_df$Var03)
group_S06_df$Var05=Remove_Outlier(group_S06_df$Var05)
group_S06_df$Var07=Remove_Outlier(group_S06_df$Var07)

```

Lets plot again plots for the variables to check if the outliers are fixed

```{r}
p1 <- ggplot(project_in_df, aes(SeriesInd, y = Var02, color = variable)) + 
    geom_line(aes(y = Var02, col = "Var02"))

p2 <- ggplot(group_S01_df, aes(SeriesInd, y = Var01, color = variable)) + 
    ggtitle("S01") + 
    geom_line(aes(y = Var01, col = "Var01"))

p3 <- ggplot(group_S02_df, aes(SeriesInd, y = Var03,color = variable))+
    ggtitle("S02") + 
    geom_line(aes(y = Var03, col = "Var03"))

p4 <- ggplot(group_S03_df, aes(SeriesInd, y = Var05, color = variable)) + 
    geom_line(aes(y = Var05, col = "Var05")) +
    ggtitle("S03") 

p4 <- ggplot(group_S03_df, aes(SeriesInd, y = Var07, color = variable)) + 
    geom_line(aes(y = Var07, col = "Var07")) +
    ggtitle("S03") 
  
p5 <- ggplot(group_S04_df, aes(SeriesInd, y = Var01, color = variable)) + 
    geom_line(aes(y = Var01, col = "Var01")) + 
    ggtitle("S04") 

p6 <- ggplot(group_S05_df, aes(SeriesInd, y = Var03, color = variable)) + 
   ggtitle("S05") +
    geom_line(aes(y = Var03, col = "Var03"))

p7 <- ggplot(group_S06_df, aes(SeriesInd, y = Var05, color = variable)) + 
    geom_line(aes(y = Var05, col = "Var05")) +
    ggtitle("S06")

p8 <- ggplot(group_S06_df, aes(SeriesInd, y = Var07, color = variable)) + 
     ggtitle("S06") +
    geom_line(aes(y = Var07, col = "Var07"))
grid.arrange(p1, p2, nrow = 1)
grid.arrange(p3, p4, nrow = 1)
grid.arrange(p5, p6, nrow = 1)
grid.arrange(p7,p8, nrow = 1)
```

The plots looks good to goahead with data

Lets create time
```{r,eval=TRUE}

```


#### <span style="color:red"><b><u>Cleanup Dataset</u></b></span>


#### <span style="color:red"><b><u>Build Model</u></b></span>


#### <span style="color:red"><b><u>Evaluate Model</u></b></span>


#### <span style="color:red"><b><u>Group S01 Forecast</u></b></span>
#### <span style="color:red"><b><u>Group S02 Forecast</u></b></span>
#### <span style="color:red"><b><u>Group S03 Forecast</u></b></span>
#### <span style="color:red"><b><u>Group S04 Forecast</u></b></span>
#### <span style="color:red"><b><u>Group S05 Forecast</u></b></span>
#### <span style="color:red"><b><u>Group S06 Forecast</u></b></span>
